---
title: "Grouping and summarizing"
format: html
editor: source
date: "2024-10-14"
output: html_document
---


### Identify the duplicated row
Identify the duplicated subject.
```{r}
# Write your answer here
library(tidyverse)
judge_url <- 
  "https://biostat2.uni.lu/practicals/data/judgments.tsv"
judgments <- readr::read_tsv(judge_url)
duplicated_subjects <- judgments%>%
  group_by(subject)|>
  filter(n()>1)
print(duplicated_subjects)
```

### Compute basic statistics for all moral dilemma columns. 
Compute the mean, the median, the standard deviation as well as min and max values.
Find meaningful short names for the functions such as `med` for  the median.

Assign `judgments_condition_stats` to the results.

```{r}
# Write your answer here
judgments_condition_stats <- judgments%>%
  summarise(across(contains("moral_dilemma"), 
                   list(
                     mean = ~ mean(.x, na.rm = TRUE),
                     med = ~ median(.x, na.rm = TRUE),
                     sd = ~ sd(.x, na.rm = TRUE),
                     min = ~ min(.x, na.rm = TRUE),
                     max = ~ max(.x, na.rm = TRUE)
                   ), .names = "{col}_{fn}"))
print(judgments_condition_stats)
```



### Sort by groups

Find the number of subjects by `age`, `gender` and `condition`, e.g. how many 20 years of age females are in the `stress` group. 

Sort the resulting tibble such that the condition that contains the most populous group is sorted first (i.e. `stress` or `control` appear together). 

```{r}
# Write your answer here
judgments|>
  group_by(age,gender,condition)|>
  count()|>
  arrange(desc(n))
```

Ensure that the resulting tibble does not contain groups.




