  ---
title: "Functional Programming"
subtitle: ""
author: "Vani Pant"
date: "28 Oct. 2024"
---

```{r setup}
#| include: false
```

> This little tutorial aims to make you familiar with some of the functions of the `stringr` package and a few regular expressions. 

## Aims 
This practical support the inline exercises of the [Functional Programming lecture](/slides/fun_programming.qmd)

## Data types reprise

#### Working with months of the year
Find three ways of expressing the names of the twelve months that will sort correctly. You can abbreviate the names. 

```{r}
#Answer1
months_str <- c(
    "January",
    "February",
    "March",
    "April",
    "May",
    "June",
    "July",
    "August",
    "September",
    "October",
    "November",
    "December"
)
factor(months_str, levels=months_str, ordered=TRUE)
```

```{r}
#Answer2
month.name
```

```{r}
#Answer3
sort(lubridate::month(1:12, label=TRUE))
```


## Functions


### What is not a function

Which programming structures do you know what are similar but not identical to _functions_?

How do they differ from _functions_?

> Your solution in text

### Calculate the median for all columns in the `trees` data set

You should be able to find three solutions, one returning a list, a vector of doubles and a tibble with a single row respectively.
```{r}
# Your solution
print(map(trees, median))
print(map_dbl(trees, median))
print(map_dfr(trees, median))


as_tibble(trees) |> 
    summarise_all(median)
```


## Build linear models

`ChickenWeights` is an in-built data set relating the growth of chicken on a diet.

Which diet has the biggest increase in growth as expressed in weight over time?



1. Split the chicken by diet using `group_split()`
2. Build linear model with the `lm` function.
3. Retrieve the model details with `summary()`. 
4. Extract the "coefficients" of the linear model which is a Matrix including a "weight" row and a "Estimate" column. 

```{r}
# Your solution
splitChick <- group_split(ChickWeight, Diet)

lm <- lm(weight ~ Diet, data = ChickWeight)
lm

lmSummary <- summary(lm)
lmSummary

ChickWeight |> group_split(Diet) |> 
    map(\(grp) lm(weight ~ Time, data=grp)) |> 
    map(summary) |> 
    map(\(x) x$coefficients["Time","Estimate"])
```




