---
title: "LEGO Project"
author: "Vani Pant"
date: "Today"
format: "html"
---


# 1. Aim
The project applies concepts from R and tidyverse lectures to analyze a dataset of official Lego sets from rebrickable.com. It focuses on creating a structured Quarto document and implementing the necessary infrastructure for effective data exploration and presentation.


# 2. The Install Script
```{r}
urls <- c(
"https://cdn.rebrickable.com/media/downloads/themes.csv.gz?1731740880.0790951",
"https://cdn.rebrickable.com/media/downloads/colors.csv.gz?1731740880.0910954",
"https://cdn.rebrickable.com/media/downloads/part_categories.csv.gz?1731740880.0990956",
"https://cdn.rebrickable.com/media/downloads/parts.csv.gz?1731740881.3191266",
"https://cdn.rebrickable.com/media/downloads/part_relationships.csv.gz?1731740909.8238513",
"https://cdn.rebrickable.com/media/downloads/elements.csv.gz?1731740882.24315",
"https://cdn.rebrickable.com/media/downloads/sets.csv.gz?1731740883.5191824",
"https://cdn.rebrickable.com/media/downloads/minifigs.csv.gz?1731740883.855191",
"https://cdn.rebrickable.com/media/downloads/inventories.csv.gz?1731740882.943168",
"https://cdn.rebrickable.com/media/downloads/inventory_parts.csv.gz?1731740908.7798247",
"https://cdn.rebrickable.com/media/downloads/inventory_sets.csv.gz?1731740908.9638295",
"https://cdn.rebrickable.com/media/downloads/inventory_minifigs.csv.gz?1731740909.3798401")

file_names <- c("themes.csv.gz", "colors.csv.gz","part_categories.csv.gz", "parts.csv.gz", "part_relationships.csv.gz", "elements.csv.gz", "sets.csv.gz", "minifigs.csv.gz", "inventories.csv.gz", "inventory_parts.csv.gz", "inventory_sets.csv.gz", "inventory_minifigs.csv.gz")

if (!dir.exists("data")){
  dir.create("data")
  }
  
for (i in seq_along(urls)) {
  file_path <- file.path("data", file_names[i])
  if (!file.exists(file_path)) {
    message(paste("Downloading:", file_names[i]))
    download.file(urls[i], file_path, mode = "wb")
    message(paste("Saved to:", file_path))
  } else {
    message(paste("Skipping:", file_names[i], "(already exists)"))
  }
}

library(tidyverse)
library(dplyr)

data_dir <- "data/"

themes <- read_csv(file.path(data_dir, "themes.csv.gz"))
colors <- read_csv(file.path(data_dir, "colors.csv.gz"))
part_categories <- read_csv(file.path(data_dir, "part_categories.csv.gz"))
parts <- read_csv(file.path(data_dir, "parts.csv.gz"))
part_relationships <- read_csv(file.path(data_dir, "part_relationships.csv.gz"))
elements <- read_csv(file.path(data_dir, "elements.csv.gz"))
sets <- read_csv(file.path(data_dir, "sets.csv.gz"))
minifigs <- read_csv(file.path(data_dir, "minifigs.csv.gz"))
inventories <- read_csv(file.path(data_dir, "inventories.csv.gz"))
inventory_parts <- read_csv(file.path(data_dir, "inventory_parts.csv.gz"))
inventory_sets <- read_csv(file.path(data_dir, "inventory_sets.csv.gz"))
inventory_minifigs <- read_csv(file.path(data_dir, "inventory_minifigs.csv.gz"))

lego_data <- read.csv("data/sets.csv.gz")
head(lego_data)
```


# 3. Warmup Questions

## 3.1. Unique Parts
```{r}
number_unique_parts <- n_distinct(parts$part_num)
parts
```


## 3.2. Colors
```{r}
number_colors <- n_distinct(colors$id)
colors
```


## 3.3. Most Frequent Colors
```{r}
color_frequency <- inventory_parts |>
  inner_join(colors, by = c("color_id" = "id")) |>
  group_by(name) |>
  summarise(count = n()) |>
  arrange(desc(count))

head(color_frequency, 10)
```


## 3.4. Top 10 Most Common Lego Parts
```{r}
top_parts <- inventory_parts |>
  group_by(part_num) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  head(10)

top_parts_name <- top_parts |>
inner_join(parts, by = "part_num")

top_parts_name
```


## 3.5. Average Number of Parts per Set
```{r}
parts_per_set <- inventory_parts |>
  group_by(inventory_id) |>
  summarise(parts_count = n())

mean_parts <- mean(parts_per_set$parts_count)
median_parts <- median(parts_per_set$parts_count)

parts_per_set
mean_parts
median_parts
```


## 3.6. The Year in which Most Lego Sets Released
```{r}
most_sets_year <- sets |>
  group_by(year) |>
  summarise(set_count = n()) |>
  arrange(desc(set_count)) |>
  slice(1)

most_sets_year
```


## 3.7. The Theme that has Most Sets Released
```{r}
most_common_theme <- sets |>
  inner_join(themes, by = c("theme_id" = "id")) |>
  group_by(theme_name =name.y) |>
  summarise(set_count = n()) |>
  arrange(desc(set_count)) |>
  slice(1)

most_common_theme
```


## 3.8. Colors by Category
```{r}
inventory_parts_with_parts <- inventory_parts |>
  left_join(parts, by = "part_num")

inventory_parts_with_categories <- inventory_parts_with_parts |>
  left_join(part_categories, by = c("part_cat_id" = "id"), suffix = c("_part", "_category"))

inventory_parts_full <- inventory_parts_with_categories |>
  left_join(colors, by = c("color_id" = "id"))

inventory_parts_no_minifigs <- inventory_parts_full |>
  filter(name_category != "Minifigs")
  
category_color_counts <- inventory_parts_no_minifigs |>
  group_by(name_category, name) |>
  summarise(count = n(), .groups = "drop") |>
  ungroup() |>
  arrange(name_category, desc(count)) |>
  pivot_wider(names_from = name, values_from = count, values_fill = 0)

inventory_parts_with_parts
inventory_parts_with_categories
inventory_parts_full
inventory_parts_no_minifigs
category_color_counts
```


# 4. Speed of Filtering Joints

## 4.1. Sampling Rows with Replacement
```{r}
library(dplyr)
library(bench)

sample_inventory_parts <- inventory_parts |>
  slice_sample(n = 10000, replace = TRUE)

sample_inventory_parts
```


## 4.2. Filtering
```{r}
filter_inventories <- sample(unique(inventory_parts$inventory_id), 100)
```


## 4.3. Filtering and Mutating Joins
```{r}
join_table <- data.frame(inventory_id = filter_inventories)

time_filter <- system.time({
  filtered_data <- sample_inventory_parts |>
    filter(inventory_id %in% filter_inventories)
})

time_semi_join <- system.time({
  semi_join_data <- sample_inventory_parts |>
    semi_join(join_table, by = "inventory_id")
})
time_inner_join <- system.time({
  inner_join_data <- sample_inventory_parts |>
    inner_join(join_table, by = "inventory_id")
})
```


## 4.4. Compare the Time for the Computation
```{r}
benchmark_results <- data.frame(
  method = c("filter", "semi_join", "inner_join"),
  user_time = c(time_filter["user.self"], time_semi_join["user.self"], time_inner_join["user.self"]),
  system_time = c(time_filter["sys.self"], time_semi_join["sys.self"], time_inner_join["sys.self"]),
  time_elapsed = c(time_filter["elapsed"], time_semi_join["elapsed"], time_inner_join["elapsed"])
)
benchmark_results
```


## 4.5. Sampling Larger Number of Rows
```{r}
sample_sizes <- c(10000, 20000, 50000, 100000)

results <- list()

for (n in sample_sizes) {
  set.seed(123)
  sampled_inventory_parts <- inventory_parts |>
    slice_sample(n = n, replace = TRUE)
  
  sampled_inventory_ids <- sampled_inventory_parts |>
    distinct(inventory_id)
  
  time_filter <- system.time({
    filtered_parts_filter <- inventory_parts |>
      filter(inventory_id %in% sampled_inventory_ids$inventory_id)
  })
  
  time_semi_join <- system.time({
    filtered_parts_semi_join <- inventory_parts |>
      semi_join(sampled_inventory_ids, by = "inventory_id")
  })
  
  time_inner_join <- system.time({
    filtered_parts_inner_join <- inventory_parts |>
      inner_join(sampled_inventory_ids, by = "inventory_id")
  })
  
  results[[as.character(n)]] <- data.frame(
    sample_size = n,
    method = c("filter", "semi_join", "inner_join"),
    time_user = c(time_filter["user.self"], time_semi_join["user.self"], time_inner_join["user.self"]),
    time_sys = c(time_filter["sys.self"], time_semi_join["sys.self"], time_inner_join["sys.self"]),
    time_elapsed = c(time_filter["elapsed"], time_semi_join["elapsed"], time_inner_join["elapsed"])
  )
}

benchmark_results_extended <- do.call(rbind, results)
print(benchmark_results_extended)
```


How fast is semi_join() in comparison to filter() and inner_join?

> From the results, we can conclude that filter() is the fastest method, followed by semi_join() and then inner_join().


Is it faster to filter with joins or by data masking using filter()?

> After running the commands again with larger sample sizes, we can still conclude that filtering using data masking with filter() is faster than using joins.
